Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Chinook Salmon Populations

### 2022-02-01 analysis/paper/paper.Rmd
update length freq plot

```{r bio-tbl}
bio_tbl = bio_g_summ %>%
  mutate(`Date Range` = season) %>%
  select(season,
         `Date Range`,
         name,
         cumm_g) %>%
  mutate(name = str_sub(name, 2, 3)) %>%
  rename(Season = season,
         `% Fish in Diet` = name,
         `g Consumed` = cumm_g) %>%
  mutate(`Date Range` = recode(`Date Range`,
                               Fall = "Sep 15 - Nov 30",
                               Spring = "Mar 1 - May 31")) %>%
  kable(booktabs = T,
        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the DSR fall and NRR spring Chinook Salmon emigrations. The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown. The Northern Pikeminnow start/end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
  kable_styling()

@@

```{r abund-data-lp}
LP_mod %>%
  select(event_name, species, data) %>%
  distinct() %>%
  select(-species) %>%
  unnest(data) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  rename(`Effort` = event_name) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of Northern Pikeminnow angling survey data for single-census mark-recapture estimators. M is the total number of fish caught in the first sample (week), marked and returned to the population; n is the total number of fish caught in the second sample (week), and m is the number of previously marked fish caught in the second week.") %>%
  kable_styling()

@@

```{r abund-data-schn}
N_mods %>%
  select(event_name, species, data) %>%
  distinct() %>%
  select(-species) %>%
  unnest(data) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  select(event_name, date, M, n, m, u, R) %>%
  rename(`Effort` = event_name,
         Date = date) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of Northern Pikeminnow angling survey data for multiple-census mark-recapture estimators. M is the total number of fish caught in previous samples (days), marked and returned to the population; n is the total number of fish caught in the current sample (day); m is the number of marked fish caught in that sample; u is the number of unmarked Northern Pikeminnow caught, and R is the number of marked fish returned to the population. Dates of sampling are also provided.") %>%
  kable_styling()

@@

```{r abund-tab}
all_est %>%
  mutate(across(c(N, SE, Lci, Uci),
                ~ prettyNum(round(.), big.mark = ","))) %>%
  rowwise() %>%
  mutate(CI = paste(Lci, Uci,
                    sep = " - ")) %>%
  select(-c(species, range:prop_range),
         -Lci, -Uci) %>%
  rename(`Effort` = event_name,
         Estimator = model,
         `95% Confidence Interval` = CI) %>%
         # `L CI` = Lci,
         # `U CI` = Uci) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Point estimates, standard errors, and 95% confidence intervals for each of the mark-recapture estimators and for the fall sampling efforts. Standard errors are not available for the multiple-census estimators.") %>%
  kable_styling()

```{r spring-N, eval = F}
spring_abund %>%
  rowwise() %>%
  mutate(`Avg. Fall N (CI)` = paste0(prettyNum(round(N_F), big.mark = ","),
                                     " (",
                                     paste0(prettyNum(round(c(Lci_F, Uci_F)), big.mark = ","),
                                            collapse = ", "),
                                     ")"),
         `Spring N (CI)` = paste0(prettyNum(round(N_S), big.mark = ","),
                                     " (",
                                     paste0(prettyNum(round(c(Lci_S, Uci_S)), big.mark = ","),
                                            collapse = ", "),
                                     ")")) %>%
  ungroup() %>%
  select(Species = species,
         `Avg Fall CPUE` = cpue_F,
         `Spring CPUE` = cpue_S,
         ends_with("(CI)")) %>%
  kable(booktabs = T,
        caption = "Average CPUE and abundance in the fall (95% confidence interval), and CPUE and estimates of spring abundance (95% confidence interval) for Northern Pikeminnow in Deadwater Slough.") %>%
  kable_styling()

@@

```{r gastric-tab}
gastric_tbl %>%
  rename(Effort = event_name,
         Captured = captured,
         Lavaged = lavaged,
         'Stomach Contents' = stmch_contents,
         'Fish Contents' = fish_contents) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of gastric lavage results including the total number of fish captured and gastric lavaged. For Northern Pikeminnow that were gastric lavaged, we show the number containing stomach contents, and of those, the number containing fish contents.") %>%
  kable_styling()

@@

@@

```{r adult-imp-tab, echo =TRUE}
impact_tbl = impact_df %>%
  select(season,
         perc_diet,
         p_chnk,
         N = juv_N,
         Lci = juv_Lci,
         Uci = juv_Uci) %>%
  group_by(perc_diet, p_chnk) %>%
  summarise(across(c("N", "Lci", "Uci"),sum),
            .groups = "drop") %>%
  mutate(Stage = "Juveniles") %>%
  mutate(p_chnk = p_chnk * 100) %>%
  rename(perc_chnk = p_chnk) %>%
  filter(perc_chnk %in% c(30, 50, 65))%>%

  bind_rows(impact_df %>%
              select(season,
                     perc_diet,
                     p_chnk,
                     N = adult_N,
                     Lci = adult_Lci,
                     Uci = adult_Uci) %>%
              group_by(perc_diet, p_chnk) %>%
              summarise(across(c("N", "Lci", "Uci"),
                   sum),
             .groups = "drop") %>%
             mutate(Stage = "Adult Equivalents") %>%
  mutate(p_chnk = p_chnk * 100) %>%
  rename(perc_chnk = p_chnk) %>%
  filter(perc_chnk %in% c(30, 50, 65))) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Estimates of juvenile fish consumed and adult equivalents for diet scenarios comprised of 10-90% fish. Estimates and 95% confidence intervals are calculated using Northern Pikeminnow abundance estimates of 28,144 (95% CI: 14,378 - 59,978) for fall and 14,897 (95% CI: 7,610 - 31,747) for spring. 'Adult equivalents' are estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for Chinook Salmon from McCann et al. 2019.") %>%
  kable_styling(font_size = 10)

```
\newpage

@@

```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort.", eval = FALSE}
total_cpue %>%
  ggplot(aes(x = event_name,
             y = catch_per_hr)) +
  geom_col(fill = "gray30") +
  labs(y = "Catch Per Angler Hour") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.title.y = element_text(vjust = 3),
        axis.text.y = element_text(size = 11,
                                   color = "black"))

@@

```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow using different mark-recapture estimators and for the fall sampling events."}

@@

  ggplot(aes(x = model,
             y = N,
             #color = model
             )) +
  geom_errorbar(aes(ymin = Lci,ymax = Uci, group=event_name), position = pos, width=0) + 
  geom_point(aes(shape = event_name), position = pos, size = 4) +
    #facet_wrap(~ event_name) +
  coord_cartesian(ylim = c(0, 100)) +
  scale_shape_manual(values = c(21, 19), name="Survey") +
  #scale_color_manual(values = c("grey35", "darkred"), name="Census type") +
  scale_x_discrete(labels=c(pm_est_labels))+
  labs(x = "Estimator",
       y = "Abundance (x 1,000)") +
  #scale_size(guide = "none")+
  theme(legend.position = c(.15, .85),legend.background = element_blank(),
        legend.key.size = unit(.5, "cm"),legend.box = 'horizontal',
        #axis.title.x = element_blank(),
        axis.text.x = element_text(size = 11, color = "black"),
        axis.text.y = element_text(size = 11, color = "black"),
        strip.text.x = element_text(size = 11,color = "black"))
  #scale_color_brewer(palette = "Dark2")

@@

```{r length-freq-p, fig.cap = "Length frequency histogram of Northern Pikeminnow caught using hook-and-line angling during the study."}
length_freq_p = gastric_df %>%
  ggplot(aes(length)) +
  geom_histogram(breaks = seq(150, 650, by = 50),
                 col = "black",
                 fill = "gray30",
                 alpha = .2) +
  labs(x = "Total Length (mm)",
       y = "Number of Northern Pikeminnow") +
  xlim(c(100, 700)) +
  ylim(c(0, 800)) +
  scale_x_continuous(breaks = seq(0, 650, by = 50)) +
  scale_y_continuous(breaks = seq(0, 550, by = 50)) +
  theme(axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"))

length_freq_p

>>>>

```{r bio-tbl}
bio_tbl = bio_g_summ %>%
  mutate(`Date Range` = season) %>%
  select(season,
         `Date Range`,
         name,
         cumm_g) %>%
  mutate(name = str_sub(name, 2, 3)) %>%
  rename(Season = season,
         `% Fish in Diet` = name,
         `g Consumed` = cumm_g) %>%
  mutate(`Date Range` = recode(`Date Range`,
                               Fall = "Sep 15 - Nov 30",
                               Spring = "Mar 1 - May 31")) %>%
  kable(booktabs = T,
        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the fall DSR and spring NRR Chinook Salmon emigrations. The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown. Northern Pikeminnow start and end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
  kable_styling()

@@

```{r abund-data-lp}
LP_mod %>%
  select(event_name, species, data) %>%
  distinct() %>%
  select(-species) %>%
  unnest(data) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  rename(`Survey` = event_name) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of Northern Pikeminnow angling survey data for the single-census mark-recapture estimator. M is the total number of fish caught in the first sample (week), marked and returned to the population; n is the total number of fish caught in the second sample (week), and m is the number of previously marked fish caught in the second week.") %>%
  kable_styling()

@@

```{r abund-data-schn}
N_mods %>%
  select(event_name, species, data) %>%
  distinct() %>%
  select(-species) %>%
  unnest(data) %>%
  mutate(across(event_name,
                ~ str_replace(., "_", " "))) %>%
  select(event_name, date, M, n, m, u, R) %>%
  rename(`Survey` = event_name,
         Date = date) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of Northern Pikeminnow angling survey data for standard Schnabel estimator. M is the total number of fish caught in previous samples (days), marked and returned to the population; n is the total number of fish caught in the current sample (day); m is the number of marked fish caught in that sample; u is the number of unmarked fish caught, and R is the number of marked fish returned to the population. Dates of sampling are also provided.") %>%
  kable_styling()

@@

```{r abund-tab}
all_est %>%
  mutate(across(c(N, SE, Lci, Uci),
                ~ prettyNum(round(.), big.mark = ","))) %>%
  rowwise() %>%
  mutate(CI = paste(Lci, Uci,
                    sep = " - ")) %>%
  select(-c(species, range:prop_range),
         -Lci, -Uci) %>%
  rename(`Survey` = event_name,
         Estimator = model,
         `95% CI` = CI) %>%
         # `L CI` = Lci,
         # `U CI` = Uci) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Point estimates, standard errors, and 95% confidence intervals for each of the mark-recapture estimators and for the fall sampling efforts. Standard errors are not available for the multiple-census estimators.") %>%
  kable_styling()

@@

```{r gastric-tab}
gastric_tbl %>%
  rename(Survey = event_name,
         Captured = captured,
         Lavaged = lavaged,
         'Stomach Contents' = stmch_contents,
         'Fish Contents' = fish_contents) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Summary of gastric lavage results including the total number of fish captured and gastric lavaged. For Northern Pikeminnow that were gastric lavaged, we show the number containing stomach contents, and of those, the number containing fish contents.") %>%
  kable_styling()

@@

```{r adult-imp-tab}
impact_tbl = total_impacts %>%
  select(perc_diet,
         p_chnk,
         N = juv_N,
         Lci = juv_Lci,
         Uci = juv_Uci) %>%
  mutate(p_chnk = p_chnk * 100) %>%
  rename(perc_chnk = p_chnk) %>%
  mutate(Stage = "Juveniles") %>%
  bind_rows(total_impacts %>%
              select(perc_diet,
                     p_chnk,
                     N = adult_N,
                     Lci = adult_Lci,
                     Uci = adult_Uci) %>%
              mutate(p_chnk = p_chnk * 100) %>%
              rename(perc_chnk = p_chnk) %>%
              mutate(Stage = "Adult Equivalents")) %>%
  select(Stage,
         `% Chinook` = perc_chnk,
         N,
         `95% L CI` = Lci,
         `95% U CI` = Uci) %>% 
  mutate(across(where(is.numeric), round)) %>%
  kable(booktabs = T,
        format.args = list(big.mark = ","),
        caption = "Estimates of juvenile Chinook Salmon consumed and adult equivalents for scenario where 60% of Northern Pikeminnow diet is fish, the median of scenarios evaluated. Table shows scenarios where 30%, 50%, and 65% of total fish consumed is Chinook Salmon. Estimates and 95% confidence intervals are calculated using Northern Pikeminnow abundance estimates of 27,874 (95% CI: 14,244 - 59,388) for fall and 14,799 (95% CI: 7,562 - 31,530) for spring. 'Adult equivalents' are estimated using a Lower Granite Dam smolt-to-adult rate (SAR) for upper Salmon River Chinook Salmon from McCann et al. (2019).") %>%
  kable_styling()

@@

@@

```{r cpue-fig, fig.cap = "The number of Northern Pikeminnow caught per angler effort for each sample effort.", eval = F}
total_cpue %>%
  ggplot(aes(x = event_name,
             y = catch_per_hr)) +
  geom_col(fill = "gray30") +
  labs(y = "Catch Per Angler Hour") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.title.y = element_text(vjust = 3),
        axis.text.y = element_text(size = 11,
                                   color = "black"))

@@

```{r abund-fig, fig.cap = "Estimates of abundance of Northern Pikeminnow in Deadwater Slough using different mark-recapture estimators and for the fall sampling events."}

@@

  ggplot(aes(x = model,
             y = N,
             #color = model
             )) +
  geom_errorbar(aes(ymin = Lci,ymax = Uci, group=event_name), position = pos, width=0) + 
  geom_point(aes(shape = event_name), position = pos, size = 4) +
    #facet_wrap(~ event_name) +
  coord_cartesian(ylim = c(0, 80)) +
  scale_shape_manual(values = c(21, 19), name="Survey") +
  #scale_color_manual(values = c("grey35", "darkred"), name="Census type") +
  scale_x_discrete(labels=c(pm_est_labels))+
  labs(y = "Abundance (x 1,000)") +
  #scale_size(guide = "none")+
  theme(legend.position = c(.15, .85),legend.background = element_blank(),
        legend.key.size = unit(.5, "cm"),legend.box = 'horizontal',
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 11, color = "black"),
        axis.text.y = element_text(size = 11, color = "black"),
        strip.text.x = element_text(size = 11,color = "black"))
  #scale_color_brewer(palette = "Dark2")

@@

```{r length-freq-p, fig.cap = "Length frequency histogram of Northern Pikeminnow caught using hook-and-line angling during the study."}
length_freq_p = gastric_df %>%
  select(event_name,
         length) %>%
  mutate(season = event_name) %>%
  mutate(season = recode(season,
                         `Fall 2019` = "Fall",
                         `Fall 2020` = "Fall",
                         `Spring 2021` = "Spring"))

ggplot(length_freq_p, 
       aes(x = length, 
           fill = season)) +
  geom_histogram(breaks = seq(150, 650, by = 25),
                 col = "black",
                 alpha = 0.5) +
  scale_fill_manual(name = "Season",
                    values = c("gray", "black"),
                    labels = c("Fall", "Spring")) +
  labs(x = "Total Length (mm)",
       y = "Frequency") +
  theme(axis.text.x = element_text(size = 11,
                                   color = "black"),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        legend.position = c(.15, .85))
